The Changing Shape of Empire: A GM Study of Chimú Bottles in Museum Collections

### 2022-01-28 supp/03-gm.analysis.Rmd
<edit gm>

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)
qdata <- qdata[match(dimnames(coords)[[3]],
                     rownames(qdata)),]
```

@@

## Define models

@@

### Vessel size and shape

```{r def.mod.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# size as a function of vessel form?
fit.size.type <- procD.lm(size ~ type,
                data = gdf,
                print.progress = FALSE,
                iter = 9999)

@@

# Procrustes ANOVA: do shapes differ as a functiuon of vessel form?
anova(fit.shape.type)
```

## Mean Shapes

>>>>

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)

qdata <- qdata[match(dimnames(coords)[[3]],
                     rownames(qdata)),]
```

@@

## Vessel allometry

```{r allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# allometry: shape as a function of vessel size?
fit.size<-procD.lm(shape ~ size, 
                   data = gdf, 
                   print.progress = FALSE, 
                   iter = 9999)

# Procrustes ANOVA: does shape change with size?
anova(fit.size)
```

```{r common.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# common allometries
fit.common<-procD.lm(shape ~ size + type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.common)

common.allometry <- plot(fit.size,
                         type = "regression",
                         reg.type = "RegScore",
                         predictor = log(gdf$size),
                         pch = pch.gps.type,
                         col = col.gps.type)

# common allometric component (Mitteroecker 2004)
plotAllometry(fit.size, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "CAC", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

```{r unique.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# unique allometries
fit.unique<-procD.lm(shape ~ size * type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.unique)

# do vessel forms express parallel, convergent, or divergent morphological characteristics?
plotAllometry(fit.unique, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "PredLine", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

## Vessel size and shape

@@

```{r def.mod.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# size as a function of vessel form (bottle/jar)?
fit.size.type <- procD.lm(size ~ type,
                data = gdf,
                print.progress = FALSE,
                iter = 9999)

@@

# Procrustes ANOVA: do shapes differ as a function of vessel form?
anova(fit.shape.type)
```

## Morphological disparity

```{r disparity, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# morphological disparity: do either of the groups display greater 
# size variation among individuals relative to the other group?
morphol.disparity(fit.size.type, 
                  groups = qdata$type, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)

# morphological disparity: do either of the groups display greater 
#shape variation among individuals relative to the other group?
morphol.disparity(fit.shape.type, 
                  groups = qdata$type, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)
```

## Mean shapes
